---
title: "TfhMAP_TCR"
author: "Damian Oyong"
date: "2023-11-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(Seurat)
library(scRepertoire)
library(data.table)
library(dplyr)
library(tidyr)
library(gtools)
library(ggraph)
library(purrr)
library(tibble)
library(circlize)
library(viridis)
library(ggalluvial)
library(RColorBrewer)
pacman::p_load(tidyr, dplyr, circlize)
```

## Import VDJ data and create scRepertoire df

Rename barcode column to match Seurat object data

```{r}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # This only works if you are using RStudio

file.names <- list.files(pattern = 'filtered_contig_annotations.csv', recursive = TRUE)
names <- list.files(path='VDJ_Mode')

file_seq <- 1:length(file.names)

if(any(grepl("*.csv", file.names))==TRUE){
  tfh_tcr <- file.names %>% 
    lapply(., function(i){
      fread(input = i)
    }) %>%
    setNames(names)
}


tfh_tcr <- combineTCR(tfh_tcr,
                      samples = names,
                      removeNA = TRUE,
                      removeMulti = TRUE,
                      filterMulti = TRUE)


## Now we need to edit the barcode, delete everything before "v3_"

tfh_tcr <- tfh_tcr %>% 
  map(~.x %>% mutate(barcode = gsub(".*v3_","",barcode)))

# tfh_tcr <- tfh_tcr[sort(names(tfh_tcr))]

tfh_tcr <- tfh_tcr[c("Tfh1_vdjMode_BoyleLab_vdj_ref_v3","Tfh2_vdjMode_BoyleLab_vdj_ref_v3",
                     "Tfh17_vdjMode_BoyleLab_vdj_ref_v3","PD1Neg_vdjMode_BoyleLab_vdj_ref_v3")]

tfh_tcr <- tfh_tcr %>%
  map2(., file_seq, ~mutate(.x, barcode = sub(1, .y, barcode)))

```

## Import Seurat

```{r}

## 
tfh_map <- readRDS('MapTfh_TRremoved_integrated_reclustered_210823.rds')

## Remove mito as per Megan's suggestion
tfh_map <- subset(tfh_map, subset = integratedcluster %in% c("cluster2", "cluster3", "heatshock", "IFNI", "ribosomal", "Tfh1", "Tfh1_CM", "Tfh1.17", "Tfh17", "Tfh2", "Tfreg"))


```

## Combine VDJ and Seurat

```{r}
tfh_rep <- combineExpression(tfh_tcr, tfh_map,
                             cloneCall="aa", 
                             proportion = FALSE, 
                             cloneSize=c(Single=1, Small=5, Medium=20, Large=100, Hyperexpanded=500))

tfh_rep@meta.data$integratedcluster <-  factor(tfh_rep@meta.data$integratedcluster, 
                                               levels = c("Tfh1", "Tfh1_CM", "Tfh2", "Tfh17", "Tfh1.17", "Tfreg", "IFNI", 
                                                          "heatshock", "ribosomal", "cluster2", "cluster3"))

```

## This is the default cluster colors and orders

```{r}

cluster_col <- c("Tfh1" = "#9e0142", "Tfh1_CM" = "#f4a582", "Tfh2" = "#5e4fa2", 
                 "Tfh17" = "#66c2a5", "Tfh1.17" = "#e6f598", "Tfreg" = "#74add1", 
                 "IFNI" = "#8c510a", "heatshock" = "#bf812d", "ribosomal" = "#fee090", 
                 "cluster2" = "#01665e", "cluster3" = "#b2abd2")

cluster_orders <- c("Tfh1", "Tfh1_CM", "Tfh2", "Tfh17", "Tfh1.17", "Tfreg", "IFNI", "heatshock", "ribosomal", "cluster2", "cluster3")

```

```{r}

# Label which cell contains either chain or both (Currently ignored)
# tfh_rep@meta.data <- tfh_rep@meta.data %>%
#   mutate(chain =
#            case_when(grepl("TRA", tfh_rep@meta.data$CTgene) & grepl("_NA", tfh_rep@meta.data$CTaa) ~ "Alpha",
#                      grepl("TRB", tfh_rep@meta.data$CTgene) & grepl("NA_", tfh_rep@meta.data$CTaa) ~ "Beta",
#                      grepl("TRA|TRB", tfh_rep@meta.data$CTgene) ~ "Both")
#   )

```

## How many cells:TCR
```{r}
cell.prop <- tfh_rep@meta.data %>% 
  mutate(TCR = case_when(is.na(CTaa) == TRUE ~ "No",
                         TRUE ~ "Yes"))

cell.prop <- cell.prop %>% 
  group_by(cloneSize) %>%
  summarise(n = n())
  # group_by(seurat_clusters) %>% 
  # mutate(total = sum(n)) %>% 
  # mutate(prop = n/total*100)

```

## Create a barplot of clonal frequency
```{r}
clone_order <- rev(mixedsort(unique(filter(tfh_rep@meta.data, !is.na(cloneType))$cloneType)))

ggplot(filter(tfh_rep@meta.data, !is.na(cloneType)), aes(x = cloneType))+
  geom_bar(fill="white", colour = "black")+
  geom_text(stat='count', aes(label= ..count..), vjust=-1)+
  scale_x_discrete(limits = clone_order)+
  theme_classic()+
  labs(y = "Count")+
  theme(axis.title.x = element_blank())
ggsave("Figures/Clonal_frequency.pdf")


tfh_rep@meta.data <- tfh_rep@meta.data %>% 
  mutate(cloneSize = case_when(clonalFrequency == 1 ~ "1",
                               clonalFrequency == 2 ~ "2",
                               clonalFrequency == 3 ~ "3",
                               clonalFrequency == 4 ~ "4",
                               clonalFrequency > 5 ~ ">5")) %>% 
  mutate(cloneSize = factor(cloneSize, levels = c('1','2','3','4','>5')))

ggplot(filter(tfh_rep@meta.data, !is.na(cloneSize)), 
       aes(x = integratedcluster, fill = cloneSize))+
  geom_bar(colour = "black", position = position_stack(reverse = TRUE))+
  # geom_text(stat='count', aes(label= ..count..), vjust=-1)+
  # scale_x_discrete(limits = clone_order)+
  theme_classic()+
  # scale_fill_brewer(palette = "GnBu", direction = -1)+
  scale_fill_manual(values = c("grey","#ABDDA4", "#66C2A5", "#3288BD", "#5E4FA2"))+
  labs(y = "Count")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90))
ggsave("Figures/Clonal_Freq_byAnnot.pdf")

```

```{r}
# This is if we want to get rid of singletons
tfh_rep <- subset(tfh_rep, subset = clonalFrequency != 1)
```

## Create barplot of expanded clones, colored on subsets

```{r}

bar_clone <- tfh_rep@meta.data %>% 
  filter(., !is.na(CTaa)) %>% 
  group_by(CTaa, integratedcluster, donor_info) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  group_by(CTaa) %>% 
  mutate(total_n = sum(count)) %>% 
  filter(., total_n > 1)

cl_order <- bar_clone %>% 
  ungroup() %>% 
  group_by(CTaa) %>% 
  summarise(total = sum(count)) %>% 
  arrange(desc(total))

ggplot(bar_clone, aes(x = CTaa, y = count))+
  scale_x_discrete(limits = cl_order$CTaa)+
  scale_y_continuous(breaks = seq(0,12, by = 2))+
  geom_col(aes(fill = integratedcluster))+
  scale_fill_manual(values = cluster_col)+
  theme_classic()+
  xlab("CTaa")+
  theme(axis.text.x = element_blank())
ggsave(filename = "Figures/ClonalCount_colored.pdf")

# Count how many are non duplicated
dups <- bar_clone %>%
  filter(!duplicated(CTaa) & !duplicated(CTaa, fromLast = TRUE))

```

## Create barplot for shared clones

```{r}
tfh_share <- getCirclize(tfh_rep,
                 group.by = "integratedcluster", proportion = T) %>% 
  filter(., from != to) 
  
tfh_share <- rbind(tfh_share %>% rename(to = from, from = to),
                  tfh_share) %>% 
  mutate(value = round(value*100, 1))  

tfh_share$from <- factor(tfh_share$from, levels = cluster_orders)

ggplot(filter(tfh_share, from != to & value != 0), aes(x = from, y = value, fill = to))+
  theme_classic()+
  ylab("Proportion of Clones Sharing")+
  labs(fill = "Cell Type")+
  geom_col(colour = "black")+
  scale_x_discrete(drop = T)+
  geom_text(aes(label = value), size = 3, position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = cluster_col)+
  theme(axis.title.x = element_blank())
ggsave(filename = "Figures/SharedClones.pdf")
```

## Create Heatmap of Clonal Overlap between donors

NOTE: These results are not bootstrapped. Ilka Wahl paper did bootsrapping 10,000. Need to find out more what's appropriate

```{r}
clonalOverlap(tfh_rep, cloneCall = "aa", 
              method = "jaccard", group.by = "donor_id")+
    theme(axis.title = element_blank(),
        # axis.text = element_blank(),
        axis.text.x = element_text(angle = 90),
        axis.ticks = element_blank()
        )

# meta <- tfh_rep@meta.data %>% 
#   distinct(donor_id) %>% 
#   remove_rownames()
# # meta$day = factor(meta$day, levels = c('day0','day8','day16','day36'))
# 
# donor_labs <- ggplot(meta, aes(donor_day, y=1, fill=donor_id)) + geom_tile() +
#   scale_fill_brewer(palette = 'Set1',name="Donor ID") + 
#   theme_void() 
# 
# day_labs <- ggplot(meta, aes(donor_day, y=1, fill=day)) + geom_tile() +
#   scale_fill_brewer(palette = 'Set3',name="Day") + 
#   theme_void() 
# 
# ibsm_overlap %>% insert_top(., donor_labs, height = .05) %>% 
#   insert_top(., day_labs, height = .05)

ggsave(filename = "Figures/ClonalOverlap.pdf")

```



## Create circos plot

Circos plot on TCR clones shared between cell types and day. Here, we want to ask what are the celltype identities clonal pairs and how they are related over time.

```{r}
# Create  a new column donor_clust for ChordDiagram "sector"
tfh_rep@meta.data$donor_clust <- paste(tfh_rep@meta.data$donor_info, tfh_rep@meta.data$integratedcluster, sep="_")

# Create circos df using getCirclize function
tfh_circos <- getCirclize(tfh_rep, group.by = "donor_clust") %>% 
  filter(., value!=0)

# Metadata on naming the group
df_names <- tfh_rep@meta.data %>% 
  select(integratedcluster, donor_info, donor_clust) %>% 
  distinct(donor_clust, .keep_all = TRUE) %>% 
  remove_rownames() %>% 
  arrange(., factor(integratedcluster))

# We need to match color donor_clust variables based on cell type
# Currently done manually, need to find easier way
names <- unique(c(tfh_circos$from, tfh_circos$to))

grid.col <- case_when(grepl("_Tfh1$", names) ~ "#9e0142",
                      grepl("_Tfh1_CM$", names) ~ "#f4a582",
                      grepl("Tfh2", names) ~ "#5e4fa2",
                      grepl("Tfh17", names) ~ "#66c2a5",
                      grepl("Tfh1.17", names) ~ "#e6f598",
                      grepl("Tfreg", names) ~ "#74add1",
                      grepl("IFNI", names) ~ "#8c510a",
                      grepl("heatshock", names) ~ "#bf812d",
                      grepl("ribosomal", names) ~ "#fee090",
                      grepl("_cluster2$", names) ~ "#01665e",
                      grepl("_cluster3$", names) ~ "#b2abd2",
                      TRUE ~ NA)

names(grid.col) <- names

# We want to split the circos based on donor grouping
group = structure(df_names$donor_info, names = df_names$donor_clust)
group = factor(group, levels = c("donor0", "donor1", "donor2","donor3"))

# Now plot backbone of circos

pdf("Figures/ChordDiagram.pdf")

chordDiagram(tfh_circos, 
             grid.col = grid.col,
             annotationTrack = "grid",
             preAllocateTracks = list(list(track.height = 0.075),
             #                          # list(track.height = 0.075),
                                      list(track.height = 0.001)),
             annotationTrackHeight = mm_h(5),
             group = group,
             big.gap = 20, small.gap = 1,
             direction.type = c("diffHeight", "arrows"))

# Now manually name the inner sector based on cell type
names <- data.frame(donor_clust = get.all.sector.index()) %>% 
  left_join(., df_names, by = "donor_clust")
names <- names$integratedcluster

for(i in seq_along(names)){
  si <- get.all.sector.index()[i]
  nm <- names[i]
  
    xlim = get.cell.meta.data("xlim", sector.index = si, track.index = 2)
    ylim = get.cell.meta.data("ylim", sector.index = si, track.index = 2)
    circos.text(mean(xlim), mean(ylim), nm, sector.index = si, track.index = 2, 
        facing = "outside", niceFacing = TRUE, col = "black", cex = 0.5)
}

# Now plot outer ribbon on days
# highlight.sector(filter(df_names, day == 'day0')$donor_clust, track.index = 1, col = viridis(n=4)[1], text = "Day 0", cex = 0.75, text.col = "white", niceFacing = TRUE)
# highlight.sector(filter(df_names, day == 'day8')$donor_clust, track.index = 1, col = viridis(n=4)[2], text = "Day 8", cex = 0.75, text.col = "white", niceFacing = TRUE)
# highlight.sector(filter(df_names, day == 'day16')$donor_clust, track.index = 1, col = viridis(n=4)[3], text = "Day 16", cex = 0.75, text.col = "white", niceFacing = TRUE)
# highlight.sector(filter(df_names, day == 'day36')$donor_clust, track.index = 1, col = viridis(n=4)[4], text = "Day 36", cex = 0.75, text.col = "white", niceFacing = TRUE)

highlight.sector(filter(df_names, donor_info == 'donor0')$donor_clust, track.index = 1, border = "black", col = "white", text = "Donor 0", cex = 0.75, text.col = "black", niceFacing = TRUE)
highlight.sector(filter(df_names, donor_info == 'donor1')$donor_clust, track.index = 1, border = "black", col = "white", text = "Donor 1", cex = 0.75, text.col = "black", niceFacing = TRUE)
highlight.sector(filter(df_names, donor_info == 'donor2')$donor_clust, track.index = 1, border = "black", col = "white", text = "Donor 2", cex = 0.75, text.col = "black", niceFacing = TRUE)
highlight.sector(filter(df_names, donor_info == 'donor3')$donor_clust, track.index = 1, border = "black", col = "white", text = "Donor 3", cex = 0.75, text.col = "black", niceFacing = TRUE)

circos.clear()

dev.off()

dev.set()
```



## Generate GLIPH check if they are CMV


## Clonal Heatmap Fidelity

```{r}
fidel <- tfh_rep@meta.data %>% 
  select(CTaa, integratedcluster) %>% 
  tibble::remove_rownames() %>% 
  group_by(CTaa, integratedcluster) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  group_by(CTaa) %>% 
  mutate(total = sum(n)) %>% 
  ungroup() %>% 
  mutate(prop = n/total) %>% 
  select(-c(n, total)) %>% 
  pivot_wider(names_from = CTaa, values_from = prop) %>% 
  arrange(., integratedcluster) %>% 
  column_to_rownames(., var = "integratedcluster") %>% 
  mutate_all(~replace(., is.na(.), 0)) %>%
  arrange()

count_per_clone <- tfh_rep@meta.data %>% 
  select(CTaa) %>% 
  tibble::remove_rownames() %>% 
  group_by(CTaa) %>% 
  summarise(n = n()) 

cluster_count <- tfh_rep@meta.data %>% 
  select(CTaa, integratedcluster) %>% 
  remove_rownames() %>% 
  group_by(integratedcluster) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n)*100) %>% 
  column_to_rownames(., var = "integratedcluster")

# column_sums <- t %>%
#   mutate_all(~replace(., is.na(.), 0)) %>%
#   summarise(across(everything(), sum)) %>%
#   pivot_longer(., everything(), values_to = "Sum") %>%
#   arrange(desc(Sum))
# 
# t <- t[column_sums$name]
fidel <- as.matrix(fidel)

library(ComplexHeatmap)

pdf("Figures/Heatmap_fidelity.pdf", width = 20)
Heatmap((fidel),
        name = "Clonal Proportion",
        cluster_columns = T, column_dend_side = "bottom", clustering_method_columns = "ward.D2",
        cluster_rows = F,
        show_column_names = F,
        # left_annotation = rowAnnotation(Prop = anno_barplot(
        #   cluster_count$prop, axis_param = list(direction = "reverse")),
        #                                                   annotation_name_rot = 90),
        top_annotation = HeatmapAnnotation(Count = anno_barplot(
          count_per_clone$n, axis_param = list(at = c(2, 5, 12))), 
          annotation_name_rot = 90),
        # rect_gp = gpar(col = "white", lwd = 0.5),
        width = ncol(fidel)*unit(0.05, "in"),
        height = nrow(fidel)*unit(0.6, "in"),
        col = colorRampPalette(RColorBrewer::brewer.pal(9, "Greens"))(100))  
        # Use continuous scale with 100 shades
dev.off()
dev.set()

```

